---
title: "Globant - Data Scientist Test"
author: "Tatiana León"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Statement

In this dataset you have 3 different outputs:

1. No readmission;
2. A readmission in less than 30 days (this situation is not good, because maybe your treatment was not appropriate);
3. A readmission in more than 30 days (this one is not so good as well the last one, however, the reason could be the state of the patient.

Your task is either to classify a patient-hospital outcome or to cluster them aiming at finding patterns that give a distinct insight.


```{r}
library(dplyr)
library(class)
library(pROC)
library(dummies)
library(ROSE)
library(sjPlot)
library(ggplot2)

setwd("C:/Users/Tatiana/Desktop/Programming")
```

## Reading and analyzing the data

We begin by reading and analyzing the data of the _Machine Learning UCI repository_ in this link: https://archive.ics.uci.edu/ml/datasets/Diabetes+130-US+hospitals+for+years+1999-2008#

The zip contains two datasets:

1. diabetic_data
2. IDs_mapping

```{r}
data <- read.csv("diabetic_data.csv", sep = ",")
IDsMapp <- read.csv("IDs_mapping.csv", sep = ",")
head(data)
```
Now we will get some insights of the data, for example, the features, its values, its structure and the relevant variables. 
```{r}
colnames(data)
str(data)
```
The first two columns encounter_id and patient_nbr are not relevant to this project, so we will remove them.

```{r}
data <- data %>% select(-encounter_id, -patient_nbr)
head(data)
```
Let's explore diag_1, diag_2 and diag_3 variables, as they have lots of values.

```{r}
diag1_table <- data %>% group_by(diag_1) %>% summarise(count = n(), prop = n()/nrow(data)) %>% arrange(desc(prop))
diag1_table
```

```{r}
diag2_table <- data %>% group_by(diag_2) %>% summarise(count = n(), prop = n()/nrow(data)) %>% arrange(desc(prop))
diag2_table
```


```{r}
diag3_table <- data %>% group_by(diag_3) %>% summarise(count = n(), prop = n()/nrow(data)) %>% arrange(desc(prop))
diag3_table
```

As they all have lots of values and none of them are significant, we will not add this variables to the model process.
```{r}
data <- data %>% select(-diag_1, -diag_2, -diag_3)
rm(diag1_table, diag2_table, diag3_table)
```

Let's take a second look of the data.

```{r}
sapply(data, unique)
```
We have some "?" values in the missing data, so we will replace this values with the corresponding NA

```{r}
data <-na_if(data, "?")
str(data)
```
Some of the variables aren't significant as they don't have density, so we will remove them. These variables are:
1. examide
2. citoglipton
3. acetohexamide
4. troglitazone
5. tolbutamide
6. glimepiride.pioglitazone
7. metformin.rosiglitazone
8. metformin.pioglitazone
9. glipizide.metformin
```{r}
data <- data %>% 
  select(-examide, -citoglipton, -acetohexamide, -troglitazone, -tolbutamide, -glimepiride.pioglitazone, -metformin.rosiglitazone, -metformin.pioglitazone, -glipizide.metformin)
sapply(data, unique)
```
Now we can convert the variables weight and age into numerical variables.

```{r}
data$weight <- ifelse(is.na(data$weight), "ND", data$weight)
data$weight <- as.factor(data$weight)
data$weight <- as.character(factor(data$weight,
                            levels = c("[0-25)", "[100-125)",  "[125-150)", "[150-175)", "[175-200)", "[25-50)" ,"[50-75)" , "[75-100)", ">200" , "ND"),
                            labels = c(25, 125,  150, 175, 200, 50 ,75 ,100, 200 , 0)),
                            stringsAsFactors=FALSE)
data$weight <- as.numeric(data$weight)

data$age <- ifelse(is.na(data$age), "ND", data$age)
data$age <- as.factor(data$age)
data$age <-  as.character(factor(data$age,
                          levels = c("[0-10)", "[10-20)", "[20-30)", "[30-40)", "[40-50)","[50-60)","[60-70)","[70-80)", "[80-90)", "[90-100)"),
                          labels = c(10, 20,  30,40, 50, 60 ,70 ,80, 90 ,100)),
                          stringsAsFactors=FALSE)
data$age <- as.numeric(data$age)
str(data)
```
The variables admission_type_id, discharge_disposition_id and admission_source_id, are categorical, although they seem numerical. We can see the meaning of each ID in the IDsMapp dataset.

Some of their values don't need to be included in our model. We will remove the rows with discharge_disposition_id equal to 11, 19, 20, 21, as they mean expired status.

```{r}
data <- data[data$discharge_disposition_id!=11,]
data <- data[data$discharge_disposition_id!=19,]
data <- data[data$discharge_disposition_id!=20,]
data <- data[data$discharge_disposition_id!=21,]
rm(IDsMapp)
```

Now let's define and explore our outcome vector.

```{r}
readm_t <- data %>% 
  group_by(readmitted) %>% 
  summarise(count = n(), prop = n()/nrow(data))
readm_t
```

We would want to predict the readmitted cases in less than 30 days. So we will define our outcome variable as follows:

```{r}
y <- as.numeric(as.factor(data$readmitted))
y <- ifelse(y == 1,1,0)
rm(readm_t)
```
We can see that we have less than 50% of the data corresponding to the outcome 1, this implies that later we have to do some oversampling work.

Now, we will set-up all the data variables.

```{r}
data <- data %>% select(-readmitted)

#get NAs
apply(apply(data, 2, is.na), 2, sum)
```
race, payer_code and medical_speciality are categorical variables, and they we NA values, we have to remove them by changing its value.

```{r}
data[is.na(data)] <- "ND"

#verify
apply(apply(data, 2, is.na), 2, sum)
```

Now, we will set-up the categorical and the numerical variables.

```{r}
cat_v <- colnames(data[,unlist(lapply(data,is.character))])
cat_v <- c(cat_v, "admission_type_id", "discharge_disposition_id", "admission_source_id")
cat_v

##numerical variables
num_v <- colnames(data[,unlist(lapply(data,is.numeric))])
num_v <- num_v[-c(3,4,5)]
num_v
```
From the categorical variables, we will reduce medical_specialty and payer_code.
```{r}
med_sp_t <- data %>% group_by(medical_specialty) %>% summarise(count = n(), prop = n()/nrow(data)) %>% arrange(desc(prop))
med_sp_t
```
Lets reduce the 10% of the values into the category "Other".
```{r}
med_spec_v <- c("ND", "InternalMedicine", "Emergency/Trauma", "Family/GeneralPractice", "Cardiology", "Surgery-General", "Nephrology", "Orthopedics" )
data  <-  data %>% mutate(medical_specialty = if_else(medical_specialty %in% med_spec_v, medical_specialty, "Other"))
rm(med_spec_v, med_sp_t)
```
We will apply the same treatment to the payer_code variable.
```{r}
payer_t <- data %>% group_by(payer_code) %>% summarise(count = n(), prop = n()/nrow(data))%>% arrange(desc(prop))
payer_t
```
```{r}
payer_v <- c("CP", "UN", "CM", "OG", "PO", "DM", "CH", "WC", "OT", "MP", "SI", "FR")
data  <-  data %>% mutate(payer_code = if_else(payer_code %in% payer_v, payer_code, "Other"))
rm(payer_v, payer_t)
```

Now we will convert the categorical variables in dummies, and then paste them to the numerical variables. This will be our preliminary final data.

```{r}
for (i in 1:25) {
  data[cat_v][,i] <- as.factor(data[cat_v][,i])
}

data_cat<-dummy.data.frame(data[cat_v])
data_num <- data[num_v]

data.d <- cbind(data_cat, data_num) 
rm(data_cat, data_num, i, cat_v, num_v)

head(data.d)
```
## Performing the model

First, we will normalize the variables of the dataset.

```{r}
normalize <- function(x)
{
  return((x- min(x)) /(max(x)-min(x)))
}
data.n <- as.data.frame(lapply(data.d[,], normalize))
```

As we said before, this dataset is imbalanced, so we have to oversampling its outcome values to get a better performance of the model.

```{r}
data.n <- cbind(data.n,y)
data.b <- ovun.sample(y~., data.n, method = "both", p = 0.5, seed = 42)
data.b <- as.data.frame(data.b$data)
```

Now, we will sample the rows.

```{r}
set.seed(42)
rows <- sample(nrow(data.b))
data.b <- data.b[rows,]
rm(rows)

y.b <- data.b$y
data.b <- data.b %>% select(-y)
```

Our final dataset to perform the model will be data.b.

#### k-Nearest Neighbor (knn)

With a piece of dataset, we will find the optimal k-value to train the knn model.

```{r}
data_test1 <- data.b[1:100,]
data_test2 <- data.b[101:200,]
y_test1 <-y.b[1:100]
y_test2 <- y.b[101:200]
```


```{r}
v <- c(0,0,0)
for (i in 1:floor(sqrt(100114*0.7))+1) {
  knn <- knn(data_test1, data_test2, cl=y_test1, k=i, prob = TRUE) 
  accuracy = (table(knn ,y_test2)[1,1] + table(knn ,y_test2)[2,2])/sum(table(knn ,y_test2))
  r <- roc(y_test2, attributes(knn)$prob)
  v <- rbind(v, c(i, accuracy, as.numeric(r$auc)))
}
v <- as.data.frame(v)
colnames(v) <- c("k", "accuracy", "auc")
v
```


```{r}
ggplot(v,aes(x=k))+
  geom_line(aes(y=auc, col = "auc"))+
  geom_line(aes(y=accuracy, col = "accuracy"))+
  labs(title="Finding the optimal k")+
  scale_color_manual(name="", 
                     values = c("auc"="#00ba38", "accuracy"="#f8766d")) + 
  theme(panel.grid.minor = element_blank())
```

We find the optimal k value in k = 60. Now we split our final data into 70% train and 30% test and then aplly the knn model.

```{r}
rm(r, v, data_test1, data_test2, y_test1, y_test2, data.d, data.n, data)

data_train <- data.b[1:70080,]
data_test <- data.b[70081:nrow(data.b),]
y_train <-y.b[1:70080]
y_test <- y.b[70081:nrow(data.b)]

knn <- knn(data_train, data_test, cl=y_train, k=60, prob = TRUE)
```

## Evaluating the model

```{r}
roc(y_test, attributes(knn)$prob)
plot(roc(y_test, attributes(knn)$prob),
     print.thres = T,
     print.auc=T)
```


```{r}
table(knn ,y_test)
```


```{r}
accuracy = (table(knn ,y_test)[1,1] + table(knn ,y_test)[2,2])/sum(table(knn ,y_test))
accuracy
```


```{r}
recall = (table(knn ,y_test)[1,1])/(table(knn ,y_test)[1,1]+table(knn ,y_test)[2,1])
recall
```


```{r}
precision = table(knn ,y_test)[1,1] /(table(knn ,y_test)[1,1]+table(knn ,y_test)[1,2])
precision
```


```{r}
specificity = table(knn ,y_test)[2,2]/(table(knn ,y_test)[1,2]+table(knn ,y_test)[2,2])
specificity
```


```{r}
prevalence = (table(knn ,y_test)[1,1]+table(knn ,y_test)[2,1])/sum(table(knn ,y_test))
prevalence
```


